---
title: "Lobsters"
format: html
editor: visual
---

# Task 2: Lobsters

Attach the necessary libraries for the program

```{r}
library(tidyverse)
library(here)
library(janitor)
```

Read in the Lobster Data File

N/A values in this dataset are stored as either "-99999" or blank cells

janitor::clean_names() \<- changes all the column names to lower snake case

dplyr::uncount(count) \<- duplicates rows according to the weighting variable in the column "COUNT"

```{r}
lobsters <- read_csv(here("data","Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) %>%
  clean_names() %>%
  uncount(count)
```

```{r}
# Create a summary table that finds the total counts (see: n()), and mean carapace lengths of lobsters observed in the dataset by site and year

mean_size_by_site <- lobsters %>%
  group_by(site) %>%
  summarize(mean_size= mean(size_mm, na.rm = TRUE),
            n())
view(mean_size_by_site)


```

```{r}
# Create a ggplot graph of the number of total lobsters observed (y-axis) by year (x-axis) in the study, grouped (either aesthetically or by faceting) by site

lobsters_obs <- lobsters %>% 
  group_by(year) %>% 
  mutate(observed_lobsters = n())

ggplot(data = lobsters_obs, aes(x=year, y = observed_lobsters))+
  geom_point()+
  facet_wrap("site")
```

Proportion of legal lobster

```{r}
legal_lobster <- lobsters %>% 
  filter(year == "2020") %>% 
  mutate(legal_limit = case_when (
    size_mm >= 79.76 ~ "legal",
    size_mm < 79.76 ~ "illegal"
  ))

legal_sum <- legal_lobster %>% 
  group_by(site,legal_limit) %>% 
  summarize(count = n())
```

```{r}
ggplot(legal_sum, aes(x = site, y = count))+
  geom_col(aes(color = legal_limit),
           position = "fill")

```

## TASK 3

```{r}
# Create and store a subset that only contains lobsters from sites “IVEE”, “CARP” and “NAPL”. Check your output data frame to ensure that only those three sites exist.
ex_a <- lobsters %>% 
  filter(site %in% c("IVEE", "CARP","NAPL"))
# Create a subset that only contains lobsters observed in August.
ex_b <- lobsters %>% 
  filter(month == "8")

# 
ex_c <- lobsters %>% 
  filter(site == "AQUE" | size_mm > 70)

ex_d <- lobsters %>% 
  filter(!site %in% "NAPL")

ex_e <- lobsters %>%
  group_by(site) %>%
  summarize(mean_size= mean(size_mm, na.rm = TRUE),
            sd_size = sd(size_mm, na.rm = TRUE))
# view(ex_e)

ex_f <- lobsters %>%
  group_by(site,month) %>%
  summarize(max_size= max(size_mm, na.rm = TRUE))

ex_g <- lobsters %>%
  mutate(size_cm = size_mm/10)

ex_h <- lobsters %>% 
  mutate(site = str_to_lower(site))

ex_i <- lobsters %>% 
  mutate(area = is.character(area))

ex_j <- lobsters %>% 
  mutate(size_bin = case_when(
    size_mm > 70 ~ "large",
    size_mm <= 70 ~ "small"
  ))

ex_k <- lobsters %>% 
  mutate(designation = case_when(
    site == "IVEE" ~ "MPA",
    site != "IVEE" ~ "not MPA"
  ))

```
